Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm

### 2022-04-18 final_metrics_analysis.Rmd
adding in processing methods

---
title: "Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm"
author: "Elisabeth Nesmith, Eunice Kim, Emma Kornberg, Amrita Acharya, Dianne Caravela"
date: "3/8/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

@@

from aif360.sklearn.preprocessing import ReweighingMeta
#from aif360.sklearn.inprocessing import 
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta

@@

```{python EDA}
X_new_index = X_new.rename(columns={"race": "def_race"})
X_new_index = X_new_index.rename(columns={"sex": "def_sex"})
# renaming dataframe columns to avoid the ValueError of variables being a column AND index label (which is ambiguous)

X_new_index.groupby(["def_race"])["age"].median()
X_new_index.groupby(["def_race", "sex"]).size()
X_new_index.groupby(["def_race", "c_charge_degree"]).size()
X_new_index.groupby(["def_race"])["priors_count"].median()
#X_new_index.groupby(["def_race"])["juv_fel_count"].median()
#X_new_index.groupby(["def_race"])["juv_misd_count"].median()
# above two lines commented out since medians for everyone = 0

```

@@

```{python exploratory priors by race}
priors_by_race = sns.countplot(x="race", hue="priors_count", data=df_viz, palette=workshop_palette)

plt.title('Priors Count by Race')
plt.show()
```

@@

# In-Processing

@@

## Prejudice Remover
```{r}
#prej = PrejudiceRemover(eta=1.0, sensitive_attr='race', class_attr='two_year_recid', seed=1234567)
#prej.fit(train_BLD)
#y_pred_PREJ = meta.predict(test_BLD)
#y_pred_PREJ = y_pred_PREJ.labels.flatten()
```

## Meta Fair Classifier 
```{python in-processing meta}
#meta = MetaFairClassifier(tau=0.8, sensitive_attr='race', type='sr', seed=1234567)
#meta.fit(train_BLD)
#y_pred_META = meta.predict(test_BLD)
#y_pred_META = y_pred_META.labels.flatten()
```

```{python in-processing accuracy score}
# acc_AD = accuracy_score(y_test, y_pred_META)
# print(f'[ Meta-Classification] The test accuracy of the algorithm is: {acc_AD: .2%}')
```

>>>>

---
title: "Analyzing Fairness: Race, Recidivism, and the COMPAS Algorithm"
author: "Elisabeth Nesmith, Eunice Kim, Emma Kornberg, Amrita Acharya, Dianne Caravela"
date: "3/8/2022"
output: html_document
---

@@

from aif360.sklearn.preprocessing import ReweighingMeta
from aif360.sklearn.postprocessing import CalibratedEqualizedOdds, PostProcessingMeta

@@

@@

@@

# In-Processing Approaches

@@

## Adversarial Debiasing
```{python adversarial debiasing}
# from aif360.algorithms.inprocessing import AdversarialDebiasing
# import tensorflow as tf
# 
# adv_deb = AdversarialDebiasing(unpr)
# #adv_deb.fit(X_train, y_train)
# #y_pred_AD = adv_deb.predict(X_test)
# #adv_deb.sess_.close()
```

## Prejudice Remover
```{python prejudice remover}
from aif360.algorithms.inprocessing import PrejudiceRemover

prej = PrejudiceRemover(eta = 1.0, sensitive_attr='race', class_attr='two_year_recid')
prej.fit(train_BLD)
y_pred_PREJ = prej.predict(test_BLD)
y_pred_PREJ = y_pred_PREJ.labels.flatten()
```

```{python prejudice accuracy score}
acc_PREJ = accuracy_score(y_test, y_pred_PREJ)
print(f'[Prejudice Remover] The test accuracy of the algorithm is: {acc_PREJ: .2%}')
```

```{python prejudice remover confusion matrix}
cf_matrix = confusion_matrix(y_test, y_pred_PREJ)
make_confusion_matrix(cf_matrix, "[Prejudice Remover]")
plt.show()
```

```{python prejudice remover metrics per group}
metrics_per_group(y_test, y_pred_PREJ)
plt.show()
```

```{python prejudice remover statistical parity difference}
stat_par_diff_PREJ = statistical_parity_difference(y_test, y_pred_PREJ, prot_attr='race', pos_label = 0)

print(f'[Prejudice Remover] The statistical parity difference is {stat_par_diff_PREJ: .2}')
```

```{python prejudice remover equal opportunity difference}
eq_opp_diff_PREJ = equal_opportunity_difference(y_test, y_pred_PREJ, prot_attr='race', pos_label = 0)

print(f'[Prejudice Remover] The equal opportunity difference is {eq_opp_diff_PREJ: .2}')
```

```{python prejudice remover average odds difference}
avg_odds_diff_PREJ = average_odds_difference(y_test, y_pred_PREJ, prot_attr='race', pos_label = 0)

print(f'[Prejudice Remover] The average odds difference is {avg_odds_diff_PREJ: .2}')
```

### Disparate Impact Ratio
```{python prejudice remover disparate impact ratio}
disp_impact_ratio_PREJ = disparate_impact_ratio(y_test, y_pred_PREJ, prot_attr='race', priv_group = 1, pos_label = 0)

print(f'[Prejudice Remover] The disparate impact ratio is {disp_impact_ratio_PREJ: .2}')
```

### Plot Fairness Metrics (Prejudice Remover)

```{python prejudice remover fairness metrics plot}
plot_fair_metrics([stat_par_diff_PREJ, eq_opp_diff_PREJ, avg_odds_diff_PREJ, disp_impact_ratio_PREJ], 'Prejudice Remover')
plt.show()
```
